---
output:
  html_document: default
  pdf_document: default
---
---
title: "R Notebook"
author: "Kadiatu Kamara"
L
https://www.dropbox.com/s/w40rxdqwconh49j/MedianZIP.xlsx?dl=0 

```{r}
# 1 Read the data
library('readxl')
getwd()
setwd("C:/Users/Kadiatu/Downloads/syracuse/wk7")
my_Data<-read.table(file = "clipboard", sep = "\t", header = TRUE)
my_Data <-read_excel('MedianZip.xlsx')
my_Data

```

# 2 Clean up the data

```{r}
#getting data details
str(my_Data)
View(my_Data)
colnames(my_Data)
nrow(my_Data)
```

#rename the columns ("zip", "Median", "Mean", "Population")
```{r}
colnames(my_Data)<-c("zip", "Median", "Mean", "Population")
my_Data
my_Data[-1,]
```


# remove any info at the front of the file that's not needed. THis will require you to explore the data
```{r}

```

 
# FYI. You'll need to remove the commas for numbers and conver to numeric (function hints: gsub() and as.numeric() )
```{r}
my_Data$zip <-as.numeric(my_Data$zip)
my_Data$Median <-as.numeric(my_Data$Median)
my_Data$Mean <-as.numeric(my_Data$Mean)
my_Data$Population <-as.numeric(my_Data$Population)
my_Data
```
```{r}
my_Data <-ceiling(my_Data)
my_Data
my_Data <-my_Data[-1,]
my_Data
```
  

# 3 Install and load the "zipcode" package. There is a dataset in zipcode that contains the following information: "zip","city","state","latitude","longitude" You can call the data by using the following code. 

```{r}
library("zipcode")
data(zipcode)
zipcode <-(zipcode)
zipcode
```

# 4 Merge the zipcode dataframe information into one dataframe using the merge() function. FYI you'll need to clean the zipcodes in in the original data by running clean.zipcodes(zips). clean.zipcodes is a function in the zipcode package.
```{r}
my_Data$zip <-clean.zipcodes(my_Data$zip)

my_DataZipCode <-merge(my_Data, zipcode, by = "zip")
my_DataZipCode <-data.frame(my_DataZipCode)
my_DataZipCode$zip <-as.numeric(my_DataZipCode$zip)
my_DataZipCode

```

# 5 Remove Hawaii and Alaska from the dataframe you created in #4
```
```{r}
my_data_state <-(my_DataZipCode)

my_data_remove_states <-subset(my_data_state, (!my_data_state$state %in% unique(my_data_state$state[state == "HI"]) & !my_data_state$state %in% unique(my_data_state$state[state == "AK"])))

my_data_remove_states

unique(my_data_remove_states$state)

```

### Step 2: Show the income & population per state

# 1 Create a simpler dataframe,with just the average median income and the population for each state (hint: create these seperately using tapply() and then combine them using merge() )

```{r}
my_dataPop <-(my_data_remove_states)
my_dataPop

my_dataPop <-na.omit(my_dataPop)
my_dataPop

my_dataPop1 <-tapply(my_dataPop$Mean, my_dataPop$state, mean )
my_dataPop1

my_dataInc <-(my_data_remove_states)
my_dataInc

my_dataInc <-na.omit(my_dataInc)
my_dataInc

my_dataInc1 <-tapply(my_dataInc$Population, my_dataInc$state, mean )
my_dataInc1

rownames(my_dataPop1)

state_abbr <-rownames(my_dataPop1)
state_abbr

my_data_Population <-data.frame(state_abbr, my_dataPop1)
my_data_Population

my_data_Income <-data.frame(state_abbr, my_dataInc1)
my_data_Income


combined_data <- merge(my_data_Income, my_data_Population, by= "state_abbr" )
combined_data <-data.frame(combined_data)
combined_data
```

  
# 2 Add the state abbreviations and the state names as new columns  get the state name (not just the abbreviations) and make sure the state names are all lower case. You can get statenames by using calling View(state.name) its a base dataset in R
```{r}
#df <-data.frame(my_dataPop1, my_dataInc1)
#df

#colnames(df) <-c("my_dataInc1", "my_dataPop1")
#colnames(df)

##dfId <- seq.int(nrow(df))
#dfId

state_abbr <-(state.abb)
state_abbr

state_name <-tolower(state.name)
state_name

states <-data.frame(state_name, state_abbr)
states

combine_all_data <-merge(states, combined_data, by = "state_abbr")
combine_all_data
```

# 3 Show the U.S. map, representing the color with the average median income of that state
```{r}
library("ggplot2")
library('maps')
library("mapdata")
library("tidyverse")
library("ggmap")
library("mapproj")

us <- map_data("state")
us

USInc <- ggplot(combine_all_data, aes(map_id = state_name))
USInc
USInc <-USInc + geom_map(map=us, aes(fill=combine_all_data$my_dataPop1), color="black")
USInc
USInc <-USInc + expand_limits(x=us$long, y=us$lat)
USInc
USInc <-USInc + ggtitle("US Income per state")
USInc

```

# 4 Show the U.S. map, with color representing the population of the state using the function tolower()
```{r}
USPop <- ggplot(combine_all_data, aes(map_id = state_name))
USPop
USPop <-USPop + geom_map(map=us, aes(fill=combine_all_data$my_dataInc1), color="black")
USPop
USPop <-USPop + expand_limits(x=us$long, y=us$lat)
USPop
USPop <-USPop + ggtitle("US Population per state")
USPop

str(us)
```

### Step 3: Show the income per zip code
Have draw each zip code on the map, where the color of the ‘dot’ is based on the median income. To make the map look appealing, have the background of the map be black.
```{r}
USZip <- ggplot(my_DataZipCode, aes( map_id = zip)) 

#USZip <-USZip + geom_map(map=us, aes( x= my_DataZipCode$Median, y = my_DataZipCode$zip))

USZip <-USZip + expand_limits(x=us$long, y=us$lat)

USZip <-USZip+ geom_point(data = my_DataZipCode, aes(x = my_DataZipCode$longitude, y = my_DataZipCode$latitude,  color= my_DataZipCode$Median))

USZip <-USZip + ggtitle("US Income per Zip Code")
USZip


```

### Step 4: Show Zip Code Density
Now generate a different map, one where we can easily see where there are lots of zip codes, and where there are few (using the ‘stat_density2d’ function).
```{r}

USZip2 <- USZip + geom_density_2d(data = my_DataZipCode, aes(x = my_DataZipCode$longitude, y = my_DataZipCode$latitude))
USZip2
USZip2 <-USZip2 + ggtitle("US Density per Zip Code")
USZip2
```
 
### Step 5: Zoom in to the region around NYC 
Repeat steps 3 & 4, but have the image / map be of the northeast U.S. (centered around New York).
```{r}




```


